### 2022-11-07 paper-thesis/chapters/04-adding-booleans.tex
More writing

This section describes how to add Booleans to the language. We will add the type of booleans `Bool`, `true`, `false` and `if`, allowing for programs such as the following:

\begin{lstlisting}
let and = \x: Bool. \y: Bool. if x then y else false end
\end{lstlisting}

\section{Postulate and Type Assert}

>>>>

This section describes how to add Booleans to the language. We will add the type of booleans \verb|Bool|, \verb|true|, \verb|false| and \verb|if|. The \verb|if| expression is not dependent, it expects both branches to have the same type. An example of a program with booleans:

\begin{lstlisting}
let and = \x: Bool. \y: Bool. if x then y else false end
\end{lstlisting}
The constructors for the language are:

\begin{lstlisting}
BoolType : Expr
BoolFalse : Expr
BoolTrue : Expr
BoolIf : Expr * Expr * Expr -> Expr
\end{lstlisting}
Then, the rules for beta reducing the language are in figure \ref{fig:bool-rules-beta}. There is one particularly interesting case, that is how to beta-reduce if statements. Converting this rule to Statix is not entirely trivial, since you need to choose which rule to apply based on what \verb|c| evaluates to. A new rule is needed, which has 3 cases. One for if \verb|c| evaluates to true, one for if \verb|c| evaluates to false, and finally a third case for other cases (such as when \verb|c| is a variable that does not have a substitution). Remember that \verb|rebuild| is the rule introduced in chapter \ref{chap:baselang}, which takes a scoped expression and a list of arguments and converts it to an expression by adding \verb|FnDestruct|s. These rules are stated below:

\begin{lstlisting}
betaReduceHead((s, BoolIf(c, b1, b2)), t) = 
  betaReduceHeadIf(s, betaReduceHead((s, c), []), c, b1, b2, t).
betaReduceHeadIf(s, (_, BoolTrue()), _, b1, _, t) = 
  betaReduceHead((s, b1), t).
betaReduceHeadIf(s, (_, BoolFalse()), _, _, b2, t) = 
  betaReduceHead((s, b2), t).
betaReduceHeadIf(s, _, c, b1, b2, t) = 
  rebuild((s, BoolIf(c, b1, b2)), t).
\end{lstlisting}

\begin{figure}[ht]
	\begin{mathpar}
		\inferrule{
		} {
			\bhr
			{ \scope{s}{\co{BoolTrue}()} }
			{ [] }
			{ \scope{s}{\co{BoolTrue}()}}
		}

		\inferrule{
		} {
			\bhr
			{ \scope{s}{\co{BoolFalse}()} }
			{ [] }
			{ \scope{s}{\co{BoolFalse}()}}
		}

		\inferrule{
		} {
			\bhr
			{ \scope{s}{\co{BoolType}()} }
			{ [] }
			{ \scope{s}{\co{BoolType}()}}
		}

		\inferrule{
			\bhr
			{ \scope{s}{c} }
			{ [] }
			{ \scope{s'}{BoolTrue()} }
			\\\bhr
			{ \scope{s}{b1} }
			{ t }
			{ \scope{s''}{b1'} }
		} {
			\bhr
			{ \scope{s}{\co{BoolIf}(c, b1, b2)} }
			{ t }
			{ \scope{s}{b1'}}
		}

		\inferrule{
			\bhr
			{ \scope{s}{c} }
			{ [] }
			{ \scope{s'}{BoolFalse()} }
			\\\bhr
			{ \scope{s}{b2} }
			{ t }
			{ \scope{s''}{b2'} }
		} {
			\bhr
			{ \scope{s}{\co{BoolIf}(c, b1, b2)} }
			{ t }
			{ \scope{s}{b2'}}
		}
	\end{mathpar}
	\caption{Rules for beta head reducing booleans}
	\label{fig:bool-rules-beta}
\end{figure}

\section{Postulate}

Next we will add \verb|Postulate| to the language.

\section{Type Assert}

Finally, we will add \verb|TypeAssert| to the language.
